---
title: "rerand"
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# rerand

The `rerand` package provides Mahalanobis rerandomization designs and
randomization-based treatment effect estimators. The package supports both a
readable R reference engine and a C++ engine for larger simulations.

## Installation

The development version can be installed from the repository with:

```{r, eval = FALSE}
remotes::install_github("YOUR_USERNAME/rerand")
```

## Design and assignment

`rerand_design()` accepts a numeric matrix or a data frame. For data-frame
input, use a one-sided formula to select balance covariates; without a formula,
every column except `id` is used. Factors and character columns are encoded
automatically. An acceptance probability or direct threshold is required.

```{r, eval = FALSE}
library(rerand)

set.seed(123)
data <- data.frame(
  id = paste0("unit-", seq_len(100)),
  age = rnorm(100, 45, 12),
  site = sample(c("A", "B", "C"), 100, replace = TRUE)
)

design <- rerand_design(
  data,
  n_treat = 50,
  formula = ~ age + site,
  id = "id",
  accept_prob = 0.1,
  seed = 123,
  engine = "cpp"
)

design
assigned_data <- as.data.frame(design)
balance(design)
```

For repeated assignment generation, prepare a reusable specification and call
`rerand_draw()`. A direct threshold can be supplied instead of `accept_prob`.

```{r, eval = FALSE}
spec <- rerand_spec(data, n_treat = 50, formula = ~ age + site,
                    id = "id", threshold = 1.5)
draws <- rerand_draw(spec, n_draws = 20, seed = 42)
```

## Estimation

`rerand_estimate()` uses a formula and a data frame. The first right-hand-side
term is the treatment column. With `method = "dim"`, a supplied design object
provides the assignment, balance covariates, and rerandomization criterion.
With `method = "lin"`, the covariates in the formula are the adjustment set;
`Y ~ Z` is intentionally rejected because it contains no adjustment covariate.

```{r, eval = FALSE}
Y0 <- 0.1 * data$age + rnorm(100)
Y1 <- Y0 + 1
assigned_data$Y <- ifelse(assigned_data$Z == 1, Y1, Y0)

estimate <- rerand_estimate(
  Y ~ Z,
  data = assigned_data,
  design = design,
  method = "dim"
)

estimate
confint(estimate)
```

For Lin's fully interacted adjustment, select covariates directly in the
formula:

```{r, eval = FALSE}
lin_estimate <- rerand_estimate(Y ~ Z + age + site, assigned_data,
                                method = "lin")
```

Use `rerand_population_stats(cbind(Y0, Y1), design = design)` for
population-level benchmarks. `get_quantile()` uses deterministic numerical
integration by default; select `method = "simulation"` only when a simulation
approximation is desired.
